---
title: "Best GLM"
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
here::i_am("ml/bestglm.qmd")
library(here)
library(tidyverse)
```

- [Bestglm R package](http://cran.nexr.com/web/packages/bestglm/)
- [Bestglm Paper](http://cran.nexr.com/web/packages/bestglm/vignettes/bestglm.pdf)


```{r}
library(bestglm)
data(zprostate)
```

Outcome: `lpsa` = Prostate-Specific Antigen (psa)

```{r}
skimr::skim(zprostate)
```
## Regsubset

```{r}
train <- (zprostate[zprostate[, 10], ])[, -10]
X <- train[, 1:8]
y <- train[, 9]
```

```{r}
out <- summary(regsubsets(x = X, y = y, nvmax = ncol(X)))
Subsets <- out$which
RSS <- out$rss


zprostate_regsub_tbl <-
  cbind(as.data.frame(Subsets), RSS = RSS) |> 
  relocate(RSS) |> 
  mutate(across(where(is.logical), ~ifelse(.x, "*", NA)))

zprostate_regsub_tbl
```
## Best GLM

### Linear Reg

```{r}
# `y` must be the last column
Xy <- cbind(as.data.frame(X), lpsa = y)

zprostate_bestglm <- bestglm(Xy)
names(zprostate_bestglm)
```


`BestModel`: lm or glm object giving the best model

```{r}
zprostate_bestglm$BestModel
```

`BestModels`: a ð‘‡ Ã— ð‘ logical matrix showing which variables are included in the top ð‘‡ models

```{r}
zprostate_bestglm$BestModels
```

`Subsets`: a (ð‘ + 1) Ã— ð‘ logical matrix showing which variables are included for subset sizes ð‘˜ = 0, . . . , ð‘ have the smallest deviance

```{r}
zprostate_bestglm$Subsets
```
`qTable`: a table showing all possible model choices for different intervals of ð‘ž.

```{r}
zprostate_bestglm$qTable
```

### Log Reg

```{r}
library(ISLR2)
data("Default")
```


```{r}
Default_Xy <- Default |> relocate(default, .after = last_col())
```


```{r}
Default_b.glm <- bestglm(Default_Xy, family = binomial(), 
                         IC = "BIC")
```


`BestModel`: lm or glm object giving the best model

```{r}
Default_b.glm$BestModel
```

`BestModels`: a ð‘‡ Ã— ð‘ logical matrix showing which variables are included in the top ð‘‡ models


```{r}
Default_b.glm$BestModels
```

### Log Reg (Penguin)

```{r}
library(palmerpenguins)
data(penguins)
# Only work with this
penguins_Xy <- penguins |> 
  relocate(sex, .after = last_col()) |> 
  select(ends_with("_mm"), body_mass_g, sex) |> 
  mutate(across(where(is.integer), as.double)) |> 
  filter(!if_any(everything(), ~is.na(.x))) |> 
  as.data.frame() # Must

head(penguins_Xy)
```

```{r}
penguins_bestglm <- bestglm::bestglm(penguins_Xy, 
                                     family = binomial())
```

```{r}
penguins_bestglm$BestModels
```

