# Infer Package Intro


```{r setup, include=FALSE}
here::i_am("infer.qmd")
library(here)
source(here("_common.R"))
```


I will explore [`{infer}`](https://infer.tidymodels.org/index.html) Package.

```{r setup-show, warning=FALSE, message=FALSE}
library(infer)
library(dplyr)
```

## Explore Data

```{r}
glimpse(gss)
```

## Specifying Response `specify()`

Specify response and explanatory variable as formula or arguments.

### Continuous Response

`age` (num) ~ `partyid` (fct)

```{r}
gss_spec_age_partyid <- gss %>% 
  specify(age ~ partyid)

# Object Type
sloop::otype(gss_spec_age_partyid)
# Class
class(gss_spec_age_partyid)
# Print
gss_spec_age_partyid
```


### Categorical Response

specifying for inference on proportions

you will need to use the `success` argument to specify which level of your response variable is a success.

```{r}
gss %>%
  specify(response = college, success = "degree")
```


## Declare the NULL Hypothesis

declare a null hypothesis using `hypothesize()`.

`null`: "independence" or "point".

### Test Independence

If the null hypothesis is that the mean number of hours worked per week in our population is 40, we would write:

```{r}
gss %>%
  specify(college ~ partyid, success = "degree") %>%
  hypothesize(null = "independence")
```
### Test Point Estimate


```{r}
gss %>%
  specify(response = hours) %>%
  hypothesize(null = "point", mu = 40)
```

## `generate()` NULL distribution 

```{r}
set.seed(1)

gss %>%
  specify(response = hours) %>%
  hypothesize(null = "point", mu = 40) %>%
  generate(reps = 1000, type = "bootstrap")
```


## Calculate Summary Stats

find the point estimate

```{r obs_mean}
obs_mean <- gss %>%
  specify(response = hours) %>%
  calculate(stat = "mean")

obs_mean
```

generate a null distribution

```{r null_dist}
null_dist <- gss %>%
  specify(response = hours) %>%
  hypothesize(null = "point", mu = 40) %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "mean")

null_dist
```

## Visualize Null Dist

```{r}
null_dist %>%
  visualize()
```

Where does our sample’s observed statistic lie on this distribution? We can use the `obs_stat` argument to specify this.

```{r}
null_dist %>%
  visualize() +
  shade_p_value(obs_stat = obs_mean, direction = "two-sided")
```

## P-value

get a two-tailed p-value


```{r}
p_value <- null_dist %>%
  get_p_value(obs_stat = obs_mean, direction = "two-sided")

p_value
```

## Confidence Interval

```{r}
# generate a distribution like the null distribution, 
# though exclude the null hypothesis from the pipeline
boot_dist <- gss %>%
  specify(response = hours) %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "mean")

# start with the bootstrap distribution
ci <- boot_dist %>%
  # calculate the confidence interval around the point estimate
  get_confidence_interval(point_estimate = obs_mean,
                          # at the 95% confidence level
                          level = .95,
                          # using the standard error
                          type = "se")

ci
```

```{r}
boot_dist %>%
  visualize() +
  shade_confidence_interval(endpoints = ci)
```

