---
title: "Diagnosis Accuracy"
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
here::i_am("stats/dta.qmd")
library(here)
library(tidyverse)
library(titanic)
library(pROC)
```


Adaped from: [Link](https://rowannicholls.github.io/R/statistics/tutorial.html#data-manipulation)

```{r}
names(titanic_train)
```


```{r}
skimr::skim(titanic_train)
```


## Survived vs Fare: Any difference ?

Let’s ask the following question: were those people who paid more for their ticket more likely to survive? 

```{r}
ggstatsplot::ggbetweenstats(
  titanic_train,
  x = Survived,
  y = Fare
)
```


Confirm the difference in `Fare` between 2 groups.

```{r}
wilcox.test(titanic_train$Fare ~ titanic_train$Survived)
```

## DTA (Manual Way)

Now let’s ask a slightly different question: can a passenger’s fare price be used to predict where or not they survived?

### Prep Data

```{r}
titanic_sub <- titanic_train |> 
  select(Fare, Survived) |> 
  mutate(Survived_orig = ifelse(Survived == 1L, "Lived", "Died")) |> 
  mutate(Survived_pred = ifelse(Fare > 14.45, "Lived", "Died")) |> 
  mutate(across(starts_with("Survived_"), 
                ~factor(.x, levels = c("Lived", "Died"))))

head(titanic_sub)
```

### Confusion Matrix

```{r}
cm <- table(pred = titanic_sub$Survived_pred, 
      orig = titanic_sub$Survived_orig)

cm
```

### Diagnostic Accuracy

```{r}
# True Positive
(tp <- cm[1, 1])
# False Positive
(fp <- cm[1, 2])
# False Negative
(fn <- cm[2, 1])
# True Negative
(tn <- cm[2, 2])
```


```{r}
# Sense
tp / (tp + fn)
# Spec
tn / (tn + fp)
# PPV
tp / (tp + fp)
# NPV
tn / (tn + fn)
```

