---
title: "Diagnosis Accuracy"
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
here::i_am("stats/dta.qmd")
library(here)
library(tidyverse)
library(titanic)
library(pROC)
```


Adaped from: [Link](https://rowannicholls.github.io/R/statistics/tutorial.html#data-manipulation)

```{r}
names(titanic_train)
```


```{r}
skimr::skim(titanic_train)
```


## Survived vs Fare: Any difference ?

Let’s ask the following question: were those people who paid more for their ticket more likely to survive? 

```{r}
ggstatsplot::ggbetweenstats(
  titanic_train,
  x = Survived,
  y = Fare
)
```


Confirm the difference in `Fare` between 2 groups.

```{r}
wilcox.test(titanic_train$Fare ~ titanic_train$Survived)
```

## DTA (Manual Way)

Now let’s ask a slightly different question: can a passenger’s fare price be used to predict where or not they survived?

### Prep Data

```{r}
titanic_sub <- titanic_train |> 
  select(Fare, Survived) |> 
  mutate(Survived_orig = ifelse(Survived == 1L, "Lived", "Died")) |> 
  mutate(Survived_pred = ifelse(Fare > 14.45, "Lived", "Died")) |> 
  mutate(across(starts_with("Survived_"), 
                ~factor(.x, levels = c("Lived", "Died"))))

head(titanic_sub)
```

### Confusion Matrix

```{r}
cm <- table(pred = titanic_sub$Survived_pred, 
      orig = titanic_sub$Survived_orig)

cm
```

### Diagnostic Accuracy

```{r}
# True Positive
(tp <- cm[1, 1])
# False Positive
(fp <- cm[1, 2])
# False Negative
(fn <- cm[2, 1])
# True Negative
(tn <- cm[2, 2])
```


```{r}
# Sense
tp / (tp + fn)
# Spec
tn / (tn + fp)
# PPV
tp / (tp + fp)
# NPV
tn / (tn + fn)
```

## ROC Curve


```{r}
r1 <- roc(Survived_orig ~ Fare, data = titanic_sub)
r1
```

```{r}
# AUC
(auc <- auc(r1))

# Confidence Interval
(ci <- ci.auc(r1))

ci_l <- round(ci[1], 2) # Lower
ci_u <- round(ci[3], 2) # Upper
```


```{r}
head(r1$thresholds)
```

### Plot (Base R)


```{r}
plot(r1, type = "S")
```


### Plot (ggplot2)


```{r}
legend_text <- paste0(
    "AUC = ", round(auc, 2), " (95% CI = ", ci_l, " - ", ci_u, ")"
)
```


```{r}
ggroc(r1)+ 
  ggtitle("Receiver Operating Characteristic Curve") +
  geom_segment(
    aes(x = 1, xend = 0, y = 0, yend = 1), color = "grey", 
    linetype = "dashed" ) +
  scale_y_continuous(expand = c(0, 0)) + 
  scale_x_reverse(expand = c(0, 0)) + 
  annotate("text", x = 0.3, y = 0.05, label = legend_text)
```






